---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## R Markdown

I imported a detailed list of my listened tracks from "https://benjaminbenben.com/lastfm-to-csv/"
after tracking my spotify usage for ~ 1 year using LastFM. I'm interested in doing some data exploration on my spotify usage and general music listening behaviour and identifying mood patterns in my choice of music. 

Starting with importing the data and some general info on the dataset:

```{r import, include=FALSE}
setwd('C:/Users/klotz/Desktop/code/lyrics NLP')              # set working directory

tracklist <- read.csv( 'kinimod21.csv', header = FALSE )
names(tracklist) <- c( 'artist', 'album', 'title','date' )   # given headers

tracklist$artist <-  as.character(tracklist$artist)          # declare appropriate the headers
tracklist$album  <-  as.character(tracklist$album)
tracklist$title  <-  as.character(tracklist$title)

library(lubridate)

tracklist$date    <- dmy_hm(as.character(tracklist$date))    # convert the datetime into various date information formats
tracklist$month   <- month(tracklist$date)
tracklist$weekday <- wday(tracklist$date)
tracklist$time    <- hour(tracklist$date)

print('first entry:')                                        # print some info
  min(date)
print('last entry:')
  max(date)
print('approximate sample period hence')
  (max(date)-min(date))
print('no of tracks recorded')
  dim(tracklist)[1]
print('average tracks per day')  
  round(dim(tracklist)[1]/as.double((max(date)-min(date)), units='days'),2)

```

Using the various date information and the nice economist theme for ggplot2 gives a nice graphic illustration of my daily, weekly and monthly usage patterns:

```{r hourly usage, echo=FALSE}
library(ggplot2)

hourly <- ggplot(data = tracklist) +
          geom_bar( aes ( factor ( time ) ), fill = c ('#014d64','#014d64','#014d64','#014d64','#014d64','#014d64','#014d64',
                                                       '#6794a7','#6794a7','#6794a7','#6794a7','#6794a7','#6794a7','#6794a7',
                                                       '#6794a7','#6794a7','#6794a7','#6794a7','#6794a7','#6794a7','#6794a7',
                                                       '#014d64','#014d64','#014d64') ) +
          coord_polar(start = -0.13, direction = 1) +
          theme_economist(base_size = 10, base_family = "verdana",
                  horizontal = TRUE, dkpanel = FALSE) + 
          theme(axis.title.y = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), axis.text.x =             element_text(face="bold"), axis.text.y = element_blank())
```


